---
title: "Self-Assessment"
---

The following document provides output from a multiple regression. Throughout, you will find circled numbers alongside pieces of output. Please complete the following Google Form that indicates each of the items on the output that you understand.

::: {.callout-important}
Mark an X on all the numbered objects you
:::

## Part 1: The Correlation Matrix (in Graphical Form)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(corrr)
library(jtools)
library(ggplot2)
library(kableExtra)
out <- datasets::mtcars %>% 
  correlate() %>% 
  rearrange() %>%
  shave()
rplot(out)

```

## Part 2: The Regression Table

```{r, echo=FALSE, message=FALSE, warning=FALSE}
mtcars$gear <- factor(mtcars$gear, labels = c("3", "4", "5"))
lm1 <- lm(mpg ~ disp + cyl + cyl:vs, mtcars)
#tbl_regression(model)
summ(lm1)
```

## Part 3: The lm() Diagnositcs


```{r, echo=FALSE, message=FALSE, warning=FALSE}
par(mfrow=c(2,2))
plot(lm1)
par(mfrow=c(1,1))
```
## Part 4: The ANOVA Table

```{r, echo=FALSE, message=FALSE, warning=FALSE}
a1 <- aov(lm1)
summary(a1)
```



