# Week 1

A new beginning for most of you.

## What do I mean?

I mean that you are starting a new course, and you are starting to learn a new programming language. You are starting to learn a new way of thinking about problems. You are starting to learn a new way of thinking about solutions. You are starting to learn a new way of thinking about the world.

```{r warning=FALSE, message=FALSE}
## a static implementation that is typical in today's teaching world
dat <- mtcars

library(ggplot2)
ggplot(dat, aes(x = wt, y = mpg)) +
  geom_point() +
  labs(
    y = "Miles per gallon",
    x = "Car's weight (1000 lbs)"
  ) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_smooth() +
  theme_minimal()
```

## Surely there is something more to this

Yes, there is. See the [next example](https://www.kaggle.com/datasets/spittman1248/cdc-data-nutrition-physical-activity-obesity/data):

```{r warning=FALSE, message=FALSE}
dat2 <- read.csv("./Nutrition__Physical_Activity__and_Obesity_-_Behavioral_Risk_Factor_Surveillance_System.csv", T)
knitr::kable(head(dat2))

```
