---
title: "Weeks 5-10"
---

```{r LoadLibraries}
#| title: "Load Libraries"
#| messages: false
#| echo: true
#| error: false
#| warning: false
#| results: "hide"
library(knitr)
library(gtsummary)
library(tidyverse)
library(flextable)
library(kableExtra)
library(ggplot2)
```



# Week 5

## Shoring up our understanding of the basics

Take the self-assessment please.

# Week 6

```{r OurData}
#| title: "Our Data"
#| messages: false
#| echo: true
#| error: false
data(mtcars)
str(mtcars)
mtcars2 <- within(mtcars, {
   vs.f <- factor(vs, labels = c("V", "S"))
   am.f <- factor(am, labels = c("automatic", "manual"))
   gear.f <- factor(gear, labels = c("3g", "4g", "5g"))
   cyl.f <- factor(cyl, labels = c("4c", "6c", "8c"))
   carb.f <- factor(carb, labels = c("1b", "2b", "3b", "4b", "6b", "8b"))
   cyl.fo  <- ordered(cyl)
   gear.fo <- ordered(gear)
   carb.fo <- ordered(carb)
})
knitr::kable(str(mtcars2))
```


## Categorical Predictors

This week (2/19) is our week to get cracking on categorical predictors. We'll start with a review of the basics, and then address two:
- Dummy coding
- Effects coding

Next week, we tackle contrast coding and polynomial coding.

### Some Basics

Categorical predictors are typically non-numerical.  In R, we refer to these predictors as factors.  Factors are a type of variable that takes on a limited number of different values; each value is referred to as a level.  For example, the variable `vs` in the `mtcars` dataset is a factor with two levels: `V` and `S`.  The variable `am` in the `mtcars` dataset is a factor with two levels: `automatic` and `manual`.  I created factors from two other variables - gear and cyl - and gave them labels that are more meaningful than the numbers that they represent.  All of these variables contain categorical data and, as a result, need to be converted into numbers so we can use them in the General Linear Model.


#### Levels, Tables, and Contrasts

You have two tools at your disposal to understand the $k$-levels of a factor: `levels()` and `table()`.  The `levels()` function returns the levels of a vector that has been stored as a factor (i.e., `as.factor()` or `factor()`.  The `table()` function returns the frequency of each level of a factor.  Additionally, the `contrasts()` function returns the contrasts used to code the levels of a factor.  We will use these functions to understand the levels of the factors we created from the `mtcars` dataset.

```{r CodingBasics}
#| title: "Coding Basics"
#| messages: false
#| echo: true
#| error: false
#| warning: false
levels(mtcars2$vs)
table(mtcars2$vs)
levels(mtcars2$vs.f)
levels(mtcars2$am)
table(mtcars2$am)
levels(mtcars2$am.f)
levels(mtcars2$gear.f)
levels(mtcars2$cyl.f)
contrasts(mtcars2$vs.f)
contrasts(mtcars2$gear.f)
contrasts(mtcars2$cyl.f)
```


### Dummy Coding

R defaults to dummy coding so if you prefer this method, just leave the handling of factors to R.  **NOT RECOMMENDED**.  Usually, we want to use a different coding method but for now, we will keep to dummy codes or, as referred to in the R world - treatment contrasts.

#### Simple 2-level Factor


```{r DummyCoding2lvl}
#| messages: false
#| echo: true
#| error: false
#| warning: false

# Base R way
contrasts(mtcars2$vs.f)
fit <- lm(mpg ~ vs.f, data = mtcars2)
summary(fit)

# Tidyverse way
mtcars2 %>%
  dplyr::as_tibble() %>%
  lm(mpg ~ vs.f, 
     data = .,
     contrasts = list(vs.f = contr.treatment)) %>%
  summary()

# means by vs.f
mtcars2 %>%
  group_by(vs.f) %>%
  summarise(mean_mpg = mean(mpg)) %>%
  kable(digits = 3)

```

Note how the `V` level of the `vs.f` factor is the reference level.  The coefficient for `vs.fS` is the difference between the mean of `mpg` for the `S` level of `vs.f` and the mean of `mpg` for the `V` level of `vs.f`.  The coefficient for `vs.fV` is actually the intercept.  So, the intercept is the reference level mean and the regression coefficient for each level of the factor is the **difference between** that reference mean.  Pretty easy, huh?

#### Simple 3-level Factor

```{r DummyCoding3lvl}
#| messages: false
#| echo: true
#| error: false
#| warning: false

# Base R way
contrasts(mtcars2$gear.f)
fit <- lm(mpg ~ gear.f, data = mtcars2)
summary(fit)

# Tidyverse way
mtcars2 %>%
  dplyr::as_tibble() %>%
  lm(mpg ~ gear.f, 
     data = .,
     contrasts = list(gear.f = contr.treatment)) %>%
  summary()

# means by gear.f
mtcars2 %>%
  group_by(gear.f) %>%
  summarise(mean_mpg = mean(mpg)) %>%
  kable(digits = 3)
```

## Effects Coding

```{r EffectsCoding2lvl}
#| title: "Effects Coding"
#| messages: false
#| echo: true
#| error: false
#| warning: false

lm(mpg ~ 1, data = mtcars2) %>%
  tbl_regression(intercept = TRUE, add_estimate_to_reference_rows=TRUE) %>% 
  add_glance_table(include = c(r.squared, adj.r.squared)) %>% 
  bold_p(t = 0.05) %>% 
  as_flex_table()

lm(mpg ~ vs.f, data = mtcars2) %>%
  tbl_regression()

lm2 <- lm(mpg ~ am.f, data = mtcars2) %>%
  tbl_regression()

lm3 <- lm(mpg ~ gear.f, data = mtcars2) %>%
  tbl_regression()

lm4 <- lm(mpg ~ cyl.f, data = mtcars2) %>%
  tbl_regression()

```



### Effects Coding


# Week 7

## Categorical Predictors (cont.)

### Contrast Coding

### Polynomial Coding

# Week 8 - SPRING BREAK (No Class)

# Week 9

## Interactions (aka Moderation)

### Simple Interactions

### Complex Interactions

# Week 10

## Moderation (cont.)

### Interactions with Categorical Predictors

### Interactions with Multiple Predictors

### Interactions with Polynomial Predictors

### Interactions with Higher-Order Predictors

