---
title: "Week 12"
---

# Tuesday

## Multiple Comparisons

Probably best to get started by understanding the topic.  We will discuss the rudiments of statistical comparison and why multiple comparisons are so important.  We will also discuss the different types of multiple comparison tests and how to perform them.  For now, I recommend everyone read Scott Maxwell's article from the syllabus and then read the following articles.

- [Getting Started with Multiple Comparisons](https://aosmith.rbind.io/2019/03/25/getting-started-with-emmeans/)
- [Estimating Marginal Means in R](https://cran.r-project.org/web/packages/emmeans/vignettes/basics.html#contents)



```{r InitSetup}
#| echo: true
#| message: false
#| warning: false
#| error: false
#| results: hide

# Load the libraries
library(tidyverse)
library(tidyr)
library(tidymodels)
library(multcomp)
library(emmeans)
library(ggplot2)

# Load the data
data(mtcars)
```

 
```{r}
#| echo: true
#| message: false
#| warning: false
#| error: false

# Create a simple model
res1 <- lm(mpg ~ factor(cyl), data = mtcars)
res2 <- aov(res1)
res3 <- aov(mpg~factor(cyl), data = mtcars)


# Create the comparison
comp1 <- glht(res1)
comp2 <- glht(res2)
comp3 <- glht(res3)

# Perform the comparison
comp1
comp2
comp3

# Note the similarities?

```

```{r newEMMs}
#| echo: true
#| message: false
#| warning: false
#| error: false

# Create the emmeans
e1m <- emmeans(res1, "cyl")
# emmeans(res2, "cyl") ## This will not work
e3m <- emmeans(res3, "cyl")

e1m
e3m

```




# Thursday

Repeated Measures

## MRC

## ANOVA



